---
title: "WEATHER ANALYSIS"
output: html_document
slug: weather
---

```{r setup, include=FALSE,message=FALSE,echo=FALSE,warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EXPLORING NASA WEATHER DATASET

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(here)
```


```{r brexit,echo=FALSE,message=FALSE,warning=FALSE}
weather <- 
  read_csv("https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.csv", 
           skip = 1, 
           na = "***")
```

```{r,warning=FALSE,message=FALSE}
#we're using the libarry KableExtra to represent our data in table format

library(kableExtra)

weather_select <- weather %>% 
  select(Year:Dec)

tidyweather <- weather_select %>% 
  pivot_longer(
    names_to = "Month",
    values_to = "delta",
    cols = -Year
  )

head(tidyweather, 20) %>% 
  kbl() %>%
  kable_paper("hover", full_width = F)
```

```{r, message=FALSE, warning=FALSE}

#we are using lubridate package to extract date, month, year from our dates
library(lubridate)
library(plotly)
tidyweather <- tidyweather %>%
  mutate(date = ymd(paste(as.character(Year), Month, "1")),
         month = month(date, label=TRUE),
         year = year(date))

tidyweather_plot <- ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies"
  )

tidyweather_plot <- ggplotly(tidyweather_plot)

tidyweather_plot
```

> As seen here the delta quantifying weather anomalies has been increasing singnificantly over the years.

```{r,message=FALSE,warning=FALSE}
tidyweather_plot +
  facet_wrap(~month)
```
```{r,message=FALSE,warning=FALSE}

comparison <- tidyweather %>% 
  filter(Year>= 1881) %>%     #remove years prior to 1881
  #create new variable 'interval', and assign values based on criteria below:
  mutate(interval = case_when(
    Year %in% c(1881:1920) ~ "1881-1920",
    Year %in% c(1921:1950) ~ "1921-1950",
    Year %in% c(1951:1980) ~ "1951-1980",
    Year %in% c(1981:2010) ~ "1981-2010",
    TRUE ~ "2011-present"
  ))
```

```{r,message=FALSE,warning=FALSE}
ggplot(comparison, aes(x=delta, fill=interval))+
  geom_density(alpha=0.2) +   #density plot with tranparency set to 20%
  theme_bw() +                #theme
  labs (
    title = "Density Plot for Monthly Temperature Anomalies",
    y     = "Density"         #changing y-axis label to sentence case
  )
```

>  According to  he desity plot major delta change occures from 1951-1980 and this is majorly because there was an increase in industrial activity which required fuel. This era also wittnessed catastrophic events such as cuban missile crisis in 1962 and two energy crises in 1973 and 1979.

```{r,warning=FALSE,message=FALSE}
#creating yearly averages
average_annual_anomaly <- tidyweather %>% 
  group_by(Year) %>%   #grouping data by Year
  
  # creating summaries for mean delta 
  # use `na.rm=TRUE` to eliminate NA (not available) values 
  summarise(annual_average_delta = mean(delta, na.rm=TRUE)) 

#plotting the data:
ggplot(average_annual_anomaly, aes(x=Year, y= annual_average_delta))+
  geom_point()+
  
  #Fit the best fit line, using LOESS method
  geom_smooth() +
  
  #change to theme_bw() to have white background + black frame around plot
  theme_bw() +
  labs (
    title = "Average Yearly Anomaly",
    y     = "Average Annual Delta"
  )                         

```

